Gradual Soundness: Lessons from Static Python

### 2022-01-27 pj-7.1/paper.tex
pj: carl MI clarifications

Relative to Python, \SP{} makes a few behavioral changes to enable
sound types and strong optimizations:
class attributes must be set in an \code{\_\_init\_\_} method to
ensure predictable layouts (all \PEP{} checkers agree on this);
module-level variables cannot be mutated; and
typed classes can inherit from at most one other typed class.

@@

>>>>

Relative to Python, \SP{} makes a few behavioral changes to enable
sound types and strong optimizations:
class attributes must be set in an \code{\_\_init\_\_} method to
ensure predictable layouts (all \PEP{} checkers agree on this);
module-level variables cannot be mutated; and
multiple inheritance among typed classes (or untyped classes with
a typed ancestor) is forbidden.

@@

%% So in normal Python you will get a layout error anytime you try to inherit from
%% more than one class defining _slots_. Any SP class with instance attributes
%% implicitly defines _slots_. So it depends how you interpret "where Python
%% wouldn't." If you consider SP classes to be slots-defining (because we emit
%% bytecode such that they are, even though they don't explicitly in the source
%% code), then our behavior is just the normal Python behavior.
%% 
%% There is one edge case where we are more strict, which is a SP class with no
%% instance attributes, only methods. In this case it will not have _slots_ and
%% Python would allow it to multiply-inherit with something else that does have
%% _slots_. But because we also have the vtables problem (which normal Python does
%% not), we can't ever allow multiple inheritance from two SP classes, even if one
%% or both of them has no instance slots, and even if that multiple inheritance
%% occurs in a nonstatic module (the nonstatic subclass will still have a vtable,
%% because we may statically invoke methods on it, since it is substitutable for
%% its static parents, and that vtable will still be impossible to construct
%% correctly.) We implement this restriction in a kind of lazy way by just
%% piggybacking on Python's existing layout incompatibility check, and adding in a
%% clause that forces every SP class to always be considered as slots-having, even
%% if it doesn't actually have any instance variable slots. So in this edge case
%% only, we will error where Python wouldn't.
%% 
%% I summarized all of this in as few words as I could with "at most one parent is
%% defined in typed code or uses _slots_ to specify its members," which is what
%% the above boils down to.
%% 
%% suppose technically we are slightly more restrictive than we would have to be:
%% we currently forbid, but technically could support a case of multiple
%% inheritance where one parent is an SP class with no instance attributes, and
%% the other parent is a nonstatic class. But this seems not very useful to
%% support, and just makes the rules even more complicated.
